### 2011-02-10 paper/hyperNeatMethod.tex
Edits to results/exploration and results table

\subsection{HyperNEAT Motion Model}
\seclabel{hyperNeatMethod}

@@

The ANN configuration follows previous studies that evolved quadruped
gaits with HyperNEAT in simulation~\cite{clune2011performance,
  clune2009evolving}, but was adapted to accommodate the physical robot used in this paper. Specifically, the ANNs consists of three $3 \times 4$
Cartesian grids of nodes forming input, hidden, and output
layers (\figref{SpiderANN.jpg}). Adjacent layers were allowed to be completely connected, meaning that there
could be $(3 \times 4)^2= 288$ links in each ANN (although evolution can set weights to 0, functionally eliminating the connection). The inputs to the
substrate were the current angles of each of the 9 joints of the robot
(described below) and a sine and cosine wave (to facilitate the
production of periodic behaviors). The sine and cosine waves were the
following function of time ($t$) in milli-seconds:
$\sin(t/12)\pi$. This function produces numbers in the range [$-\pi,
  \pi$], which were then JASON FILL ME IN TO DESCRIBE HOW JOINT
COMMANDS WERE SPECIFIED.

\editbox{are we mentioning that we gave HyperNEAT the center joint
  here or earlier?}

\edit{jason: edit this section to reflect the post-processing...}

>>>>

\subsection{HyperNEAT Gait Generation and Learning}
\seclabel{hyperNeatMethod}

@@

The ANN configuration follows previous studies that evolved quadruped
gaits with HyperNEAT in simulation~\cite{clune2011performance,
  clune2009evolving}, but was adapted to accommodate the physical robot used in this paper. Specifically, the ANNs consists of three $3 \times 4$
Cartesian grids of nodes forming input, hidden, and output
layers (\figref{SpiderANN.jpg}). Adjacent layers were allowed to be completely connected, meaning that there
could be $(3 \times 4)^2= 288$ links in each ANN (although evolution can set weights to 0, functionally eliminating the connection). The inputs to the
substrate were the current angles of each of the 9 joints of the robot and a sine and cosine wave (to facilitate the
production of periodic behaviors). The sine and cosine waves were the
following function of time ($t$) in milli-seconds:
$\sin(t/12)\pi$, which produces numbers in the range [$-\pi, \pi$].

\editbox{Jeff: this can't be right, because it would imply a period of only 12*2*pi=75 ms. Perhaps this was as a function of time in units of timesteps?}
\editbox{Jeff: proofread this paragraph to make sure I didn't lie (too
  much)}

The outputs of the substrate at each time step were nine numbers in
the range $[-1,1]$, which were scaled according to the allowable
ranges for each of the nine motors and then commanded
the positions for each motor.  Occasionally HyperNEAT would produce networks that
exhibited rapid oscillatory behaviors, switching from extreme negative to extreme positive numbers each time step.  This resulted in motor
commands to alternate extremes every 25ms (given the command rate of
40Hz), which tended to damage and overheat the motors.  To ameliorate
this problem, we requested four times as many commanded
positions from HyperNEAT ANN's and averaged over four commands at a time to
obtain the actual gait $g(t)$.  This solution worked well and did not restrict the expressiveness of HyperNEAT.

%\editbox{are we mentioning that we gave HyperNEAT the center joint
%  here or earlier?}
